\documentclass{article}

\usepackage{amsmath, tikz, gensymb, cancel}
\usepackage{microtype, enumitem, float}
\usepackage[letterpaper]{geometry}

\usetikzlibrary{calc}

\newcommand{\x}{\mathrm{x}}
\newcommand{\y}{\mathrm{y}}
\newcommand{\z}{\mathrm{z}}

\title{Homework Problem \#48}
\author{Laith}
\date{2/7/2023}

\begin{document}
\maketitle

\section*{Problem}
In Fig. 4-41, a ball is thrown up onto
a roof, landing 4.00 s later at height $h=20.0$ m 
above the release level. The ball's path just before 
landing is angled at $\theta=60.0\degree$ with the 
roof. 
\begin{enumerate}[label=(\alph*)]
    \item Find the horizontal distance $d$ it travels. 
    (See the hint to Problem 39.)
    \item What is the magnitude relative to the horizontal 
    of the ball's initial velocity?
    \item What is the angle relative to the horizontal of the 
    ball's initial velocity?
\end{enumerate}

\begin{figure}[H]
    \caption{Figure 4-11}
    \begin{tikzpicture}
        \draw[dashed] (-1, 0) -- (10, 0);

        \draw[dashed, -stealth, very thick, red] (0, 0) .. controls (2, 5) and (4, 6) .. (6.5, 3);

        \draw[solid] (5, 3) -- (8, 3);
        \draw[solid] (5, 0) -- (5, 3);
        \draw[solid] (8, 0) -- (8, 3);

        \draw[stealth-] (6.5, 0) -- (6.5, 1);
        \node at (6.5, 1.5) {$h$};
        \draw[-stealth] (6.5, 2) -- (6.5, 3);

        \draw (0, -2) -- (0, -1);
        \draw (6.5, -2) -- (6.5, -1);

        \draw[stealth-]  (0, -1.5) -- (3, -1.5);
        \node at (3.25, -1.5) {$d$};
        \draw[-stealth]  (3.5, -1.5) -- (6.5, -1.5);

        \draw (6.5, 3) -- (5.9, 4);

        \draw (6.2, 3.5) arc (-270:-180:0.5) node[above left, yshift=5, xshift=2] {$\theta$};

    \end{tikzpicture} 
\end{figure}

\newpage
\section*{Solution}

\subsection*{Part (a)}
First we need to find the distance $d$. Since this distance is 
horizontal, we can define it as the function $\y(t)$. We can 
determine $\y(t)$ to be equal to $v_0\cos(\theta)t$. We already know 
that our time will be $t=0$ as that is when the ball is thrown.
However, we do not know what our initial velocity $v_0$ is. 
We also can't solve for $v_0$ since $\y(t)$ and the angle at which the 
ball was thrown are also unknown. We do know that our height at $t=4$ is 
20 m and the ball lands with an angle of $\theta=60\degree$. So, 
by defining our height as the function $\z(t)$, which will be equal to 
$v_0\sin(\theta)t-\frac{1}{2}gt^2$, in which $\z(4)$ = 0, we can solve for $v_0$ as that will be our only unknown,
and then we can substitute $v_0$ to get distance.
\begin{align}
    &\z(t) = v_0\sin(\theta)t -\frac{1}{2}gt^2\\
    &\z(4) = v_0\sin(60)(4) -\cancelto{5}{\frac{1}{2}(10)}(4)^2\\
    &h = v_0\frac{\sqrt3}{\cancelto{1}{2}}(\cancelto{2}{4}) - 5(16) \\
    &20 = v_02\sqrt3 - 80 \Rightarrow 100 = v_02\sqrt3\\
    &\Rightarrow v_ 0 = \frac{10\sqrt3}{3} = \frac{10}{3}\sqrt3
\end{align}
We find that our initial velocity is $\frac{10}{3}\sqrt3$ m/s and we can now substitute 
this back into our equation for distance:
\begin{align}
    &\y(t) = v_0\cos(\theta)t \\
    &\y(4) = (\frac{10}{3}\sqrt3)\cos(60)(4) \\
    &d = (\frac{10}{3}\sqrt3)\cancelto{2}{(\frac{1}{2})(4)} \\
    &d = 2(\frac{10}{3}\sqrt3) = \frac{20}{3}\sqrt3
\end{align}
Our answer:
\[\boxed{d = \frac{20}{3}\sqrt3\,\mathrm{m}}\]

\end{document}